\chapter{Control Architecture}
\section{Introduction}
Vukobratovic \cite{Vuk1970} was one of the first researchers involved in the stability of bipedal robots, followed by Kajita \cite{Kaj2001} and Kim\cite{Kim2004}. In all their studies, the biped robot was usually represented by a planar inverted pendulum with the base representing the foot and the ankle joint. 
Latest researchs are focused on divide robot balance control into different strategies: ankle, hip and step strategies. The basis of the three stategies are close to the ZMP areas explained in previous sections. When the robot is in a stable posture and a disturbance is applied, depending on the magnitude or the that distrubance, the robot will react different. If the change of the ZMP position remains in a stable area, the control will react by the motion of the ankle joints to recover the robot balance. Nevertheless, if the perturbation increases and the ZMP position reaches an uncertain-stability area, it will be also necessary to move the hip joints in addition to the ankles to recover balance. Even a gait will be necessary if the loss of stability is unavoidable. 

A humanoid is an electromechanical system, so it should have all type of errors: structure flexion, small blacklashes between motion parts, etc. Also it will operate in a co-existing environment with humans, so the disturbances are unexpected at any time. Therefore, the Stabilizer is an essential element to provide stable human-like walking of a humanoid robot. The Stabilizer should perform two basic operations:
\begin{itemize}
\item[1.] When the humanoid robot walks, it should correct the robot’s walking trajectory in order to provide the secure position at any time of its motion.
\item[2.] When the humanoid robot has stopped, it should control its posture.
\end{itemize}

Thus, the Stabilizer can be decoupled into ZMP and Attitude controllers. Attitud controller is applied when velocity and accelerations of the robot boy are measured. This Master Thesis will deal with the issue of maintaining an upright posture while the robot is in a static position (without following a motion pattern) by only and moving the robot ankle joints. Then the Stalibizer designed in this work, will be responsible to control the robot ZMP position using the F-T sensors located at the robot ankles.


\input{04_control_architecture/00_inv_pendulum_model}
\input{04_control_architecture/01_LQRdesign}
\input{04_control_architecture/02_stabilizer}
\input{04_control_architecture/03_Control_strategies}

